---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Spotify200

<!-- badges: start -->
<!-- badges: end -->

Spotify200 is a data package centered on Spotify's Top 200 Most Streamed songs in the U.S. It includes a wealth of information about every song which reached the top 200 between January 1, 2017 and December 2019, including audio-features, artist-specific information, and its ranking over time.

## Installation

You can install the released version of Spotify200 with:

``` r 
devtools::install_github("IAjimi/Spotify200", build_vignettes = TRUE)
```

## Datasets

The following displays the first rows of all datasets included in this package:

```{r example}
library(Spotify200)
head(chart_position)
head(track_information)
head(track_position_aggregate)
head(pitchfork_reviews)
```

More information about the datasets can be found in the vignettes.

## Data Collection
The data collection for this package was done in three steps:

1. Information about the 200 most streamed songs per day in the U.S. was downloaded from the [Spotify Charts website](https://spotifycharts.com/). The information is available as `.csv` files available for download from the site: the download link varies predictably with the date the information was collected on, so all files can easily be retrieved with a for loop. 
2. The [Spotify API](https://developer.spotify.com/documentation/web-api/) was used to retrieve artist and track specific information for all tracks found in 1. There is a Spotify API in R, but it doesn't work properly when user authorization is required. The user authorization flow isn't actually required for this package but I still used my own API wrapper functions (customized to retrieve the information I was interested in).
3. Review information for the artists in 1) was web-scraped from [Pitchfork](https://pitchfork.com/), a well-known 'indie' music review publication.

The code can easily be extended to collect the most recent information from Spotify Charts: since the code for steps 2) and 3) is written to find information relevant to 1), the data collection in those steps automatically extends to collect this new, additional information. Similarly, the url used in 1) can easily be amended to retrieve the same information for any other country.

Generally, the code is highly dependent on web-scraping, the Spotify API, and specific URLs, so a change in any of the three would require a corresponding rewrite.
