---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Spotify200

<!-- badges: start -->
<!-- badges: end -->

Spotify200 is a data package centered on songs which reached Spotify's Top 200 Most Streamed songs in the U.S. between January 2017 and December 2019.

## Installation

You can install the released version of Spotify200 with:

``` r 
devtools::install_github("IAjimi/Spotify200")
```

## Datasets

The following displays the first rows of all datasets included in this package:

```{r example}
library(Spotify200)
head(chart_position)
head(track_information)
head(track_position_aggregate)
head(pitchfork_reviews)
```

## Data Collection
The data collection for this package was done in three steps:

1. Information was downloaded from the Spotify Charts website about the 200 most streamed songs in the U.S. The information is available as `.csv` files available for download from the site: the download link varies predictably with the date the information was collected on, so all files can easily be retrieved with a for loop. 
2. The Spotify API was used to retrieve artist and track specific information for all tracks found in 1. There is a Spotify API in R, but it doesn't work properly when user authorization is required. The user authorization flow isn't actually required for this package, but I still used my own API wrapper functions (customized to retrieve the information I was interested in).
3. Review information for the artists in 1) was web-scraped from Pitchfork.

The code can easily be extended to collect the most recent information from Spotify Charts: since the code for steps 2) and 3) is written to find information relevant to 1), the data collection in those steps automatically extends to collect this new, additional information.

Generally, the code is highly dependent on web-scraping, the Spotify API, and specific URLs, so a change in any of the three would require a corresponding rewrite.
