---
title: "chart_position"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{chart_position}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("ggplot2")
library("dplyr")
library("Spotify200")
```

## Chart Position

This vignette briefly shows some of the information one can derive from the `chart_position` dataset.

First, we can look at total daily streams streams per artist over time.

```{r}
# Total Streams per Artist per Day
time_position <- chart_position %>% 
  group_by(Artist, Date) %>%
  summarize(Streams = sum(Streams),
            per = Streams / sum(.$Streams)) %>%
  arrange(desc(Date))

# Find Top Artists
top_artists <- time_position %>% arrange(desc(Streams)) %>% distinct(Artist)

# Graph: Total Streams per Date per Artist
ggplot(time_position, aes(Date, Streams, group = Artist)) +
  geom_line(color = "grey") +
  geom_line(data = filter(time_position, Artist %in% top_artists$Artist[1:5]) , aes(color = Artist)) +
  labs(title = "Streams Per Artist Over Time", y = "Total Daily Streams")
```

We can also see who were the most streamed artists from 2017 to 2019 as well as in 2019 alone:

```{r}
## Most Streamed Artists Overall
chart_position %>% 
  group_by(Artist) %>%
  summarize(Streams = sum(Streams),
            per = Streams / sum(.$Streams)) %>%
  arrange(desc(Streams)) %>%
  head(10)

## Most Streamed Artists in 2019
chart_position %>% 
  filter(Date >= "2019-01-01") %>%
  group_by(Artist) %>%
  summarize(Streams = sum(Streams),
            per = Streams / sum(.$Streams)) %>%
  arrange(desc(Streams)) %>%
  head(10)
```

We can note, in passing, that most artists are either pop or hip-hop artists (sometimes both). Moreover, the 2019 scene seems to have been dominated by new artists, such as Billie Eilish, Juice WRLD, Lil Nas X and DaBaby.

We can look at which songs were the most streamed throughout this time period:

```{r}
## Most Streamed Songs Overall
chart_position %>% 
  group_by(Artist, Track) %>%
  summarize(Streams = sum(Streams),
            per = Streams / sum(.$Streams)) %>%
  arrange(desc(Streams)) %>%
  head(10)
```

Finally, we can find the most streamed songs per day, as well as the songs which were in the top position the longest:
```{r}
## Most Streamed Songs in A Day
chart_position %>% arrange(desc(Streams)) %>% select(Track, Artist, Date, Streams) %>% head(10)

## Songs Longest in Number 1 Position
chart_position %>% filter(Position == 1) %>% group_by(Track, Artist) %>% count() %>% arrange(desc(n)) %>% head(10)
```

Drake had the most streams per day after the release of *Scorpion*, his fifth studio album, on June 29, 2018. This success was long-lived, as Drake also had 3 of the songs which stayed in the top position for the longest.
