---
title: "pitchfork_reviews"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pitchfork_reviews}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("ggplot2")
library("dplyr")
library("Spotify200")
```

## pitchfork_reviews

This Vignette briefly shows some of the information one can derive from the `pitchfork_reviews` dataset. It is important to remember that this dataset does not include data for all Pitchfork reviews, but only the subsegment relating to artists which were listed in `chart_positions`. These reviews are all of *albums*, so the number of observations in this dataset is necessarily smaller than that of `track_list`.

First, we can look at the highest rated albums.

```{r}
# More Appropriate Variable Types
pitchfork_reviews$score <- as.numeric(pitchfork_reviews$score)
pitchfork_reviews$pub_date <- as.Date(pitchfork_reviews$pub_date)

## Highest Rated Albums
pitchfork_reviews %>% arrange(desc(score)) %>% head(10) #highest scoring albums

## Highest Rated Artists
pitchfork_reviews %>% 
  group_by(artist) %>%
  summarize(score = mean(as.numeric(score), na.rm = TRUE),
            n = n()) %>%
  filter(!is.na(score)) %>%
  arrange(desc(score))

```

