---
title: "track_information"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{track_information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Spotify200)
library(dplyr)
library(ggplot2)
library(stringr)
```

## track_information

This dataset, as this track indicates, includes a wealth of track-level information. The inclusion of popularity features (which can be complemented with the chart records included in this package) makes this a good source of information for regressions seeking to explain the popularity of certain songs.


```{r}
summary(track_information)
```

We can look at the correlation between track popularity and other key variables:
```{r}
track_information %>% 
  select_if(is.numeric) %>% 
  select(- score, - bnm, - n_albums_review) %>%
  filter(!is.na(artist_popularity) & !is.na(followers)) %>% 
  cor() %>% 
  round(2)
```

We can also look at the most popular genres:
```{r}
track_information %>% mutate(
  genres = as.character(genres), #cleaning up genres
  simple_genre = 
  case_when(
    str_detect(genres, "(hip hop|rap|grime)") == TRUE ~ "hip hop",
    str_detect(genres, "soul") == TRUE ~ "soul",
    str_detect(genres, "pop") == TRUE ~ "pop",
    str_detect(genres, "r&b") == TRUE ~ "r&b",
    str_detect(genres, "latin") == TRUE ~ "latin",
    str_detect(genres, "hollywood") == TRUE ~ "hollywood",
    str_detect(genres, "country") == TRUE ~ "country",
    str_detect(genres, "adult standards") == TRUE ~ "adult standards",
    str_detect(genres, "(dance)|(house)|(techno)") == TRUE ~ "dance",
    str_detect(genres, "rock") == TRUE ~ "rock",
    T ~ genres #else leave intact
  )
) %>%
  group_by(simple_genre) %>%
  summarise(n = n(),
            popularity = median(popularity, na.rm = TRUE)) %>%
  arrange(desc(popularity)) %>%
  filter(n > 1) #remove outlier genres
```

This isn't very accurate, as more popular genres have more songs in the top 200 and therefore have a lower average / median popularity.

We can also use this data to look at the median track features of popular artists:

```{r}
# Most Popular Artists
top_10_artists <- track_information %>% arrange(desc(artist_popularity)) %>% distinct(Artist, artist_popularity) %>% top_n(50, artist_popularity)

# Mean Values For All Top 50 Artists
track_information %>% 
  filter(Artist %in% top_10_artists$Artist) %>% 
  group_by(Artist) %>%
  select_if(is.numeric) %>%
  mutate_all(mean) %>%
  head()
``` 

A more fine-tuned analysis could look at how those metrics change over time.

